Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-08-31 fully-abstract-comp/technical-report.tex
everything so far. lots of holes here and there.

%%% Math notation
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\defbnf}{::=}
\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
\newcommand{\ssem}[2][\Phi]{\sem{#2}_{\mathrm{src}}(#1)}
\newcommand{\dom}{\mathrm{dom}}

@@

@@

@@

% One argument
\newcommand{\sjmp}[1]{\oneinstr{\src{jmp}}{#1}}
\newcommand{\ssetatob}[1]{\oneinstr{\src{seta2b}}{#1}}
\newcommand{\spush}[1]{\oneinstr{\src{push}}{#1}}
\newcommand{\spop}[1]{\oneinstr{\src{pop}}{#1}}

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{isptr}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
\newcommand{\sgetl}[2]{\twoinstr{\src{getl}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\ssubseg}[3]{\threeinstr{\src{subseg}}{#1}{#2}{#3}}
\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
\newcommand{\scseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

@@

% One argument
\newcommand{\tjmp}[1]{\oneinstr{\trg{jmp}}{#1}}
\newcommand{\tsetatob}[1]{\oneinstr{\trg{seta2b}}{#1}}

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{isptr}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
\newcommand{\tgetl}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}

% Three arguments
\newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
\newcommand{\tsubseg}[3]{\threeinstr{\trg{subseg}}{#1}{#2}{#3}}
\newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
\newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
\newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}
\newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

@@

%%% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\rn}{\var{rn}}
\newcommand{\reg}{\var{reg}}
\newcommand{\mem}{\var{mem}}
\newcommand{\ms}{\var{ms}}
\newcommand{\pc}{\var{pc}}
\newcommand{\stk}{\var{stk}}
\newcommand{\ret}{\var{ret}}
\newcommand{\data}{\var{data}}
\newcommand{\priv}{\var{priv}}
\newcommand{\opc}{\var{opc}}
\newcommand{\vsc}{\var{sc}}
\newcommand{\cb}{\var{cb}}
\newcommand{\baddr}{\var{b}}
\newcommand{\eaddr}{\var{e}}
\newcommand{\aaddr}{\var{a}}
\newcommand{\stdrng}{[\baddr,\eaddr]}

@@

%%% Functions
\newcommand{\plainfun}[2]{
  \ifthenelse{\equal{#2}{}}
  {\mathit{#1}}
  {\mathit{#1}(#2)}
}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}

\begin{document}
\section{The two capability machines}
\subsection{Domains}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    & \aendbnf & \in & \Addr \uplus \{\infty \} \\
    &  \sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rretc \mid \rretd \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

In the source language, $\src{\Stack}$ is a call stack that contains the data for each call. The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) the old data capability, 3) caller private stack, and 4) the address of the old stack pointer.

More convenient definitions
\[
  \begin{array}{rcl}
    \Phi(r) & \defeq & \Phi.\reg(r)
  \end{array}
\]
where $r\in \RegName$

\subsection{Syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (Inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
\end{array}
\]

The source machine is also a capability machine with memory capabilities, local capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rretc, \rretd, \rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

\subsubsection{Helpful functions}
\[
  \updPcAddr{} \defeq \dots
\]

@@

@@

\[
  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rretc, \rretd, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
\]

@@

If we are not in one of the two cases above, then the execution fails.
\begin{align*}
  \ssem{\scall{r_1}{r_2}} = & 
    \begin{cases}
      (\mem,\reg',\stk',\ms_\stk') & \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        % "red conditions" (reference to blackboard picture)
                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
                                        \nonExec{cb_2} \text{ and } \\
                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
                                        % "yellow conditions"
                                        \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
                                        \noRetStkReg{\reg} \text{ and } \\
                                        \noRetStkMs{\ms_\stk'} \text{ and } \\
                                        \nonLoc{\cb_1,\cb_2} \text{ and } \\
                                        \reg' = \reg
                                        \begin{array}[t]{l}
                                        \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
                                        \update{\rretc}{\retptrc}\\
                                        \update{\rretd}{\retptrd}\\
                                        \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
                                        \end{array}
 \\
                                        \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
                                        \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
                                        \stk' = (\opc,\Phi(\rdata),\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
                                        \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
                                      \end{array}\\
                                      & \\
      \undefined & \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        % "red conditions" (reference to blackboard picture)
                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
                                        \nonExec{cb_2} \text{ and } \\
                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
                                        % "yellow conditions"
                                        \text{not } \arraycolsep=0pt
                                        \left(\begin{array}{l}
                                          \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
                                          \noRetStkReg{\reg} \text{ and } \\
                                          \noRetStkMs{\ms_\stk'} \text{ and } \\
                                          \nonLoc{\cb_1,\cb_2}
                                        \end{array} \right)
                                      \end{array}\\
                                      & \\
      \failed & \text{otherwise}
    \end{cases}
\end{align*}
\lau{do we need to require anything on the data capability we put on the call stack?}

The $\scseal{}{}{}$ instruction seals a sealable capability with a seal.
\begin{align*}
  \ssem{\scseal{r_1}{r_2}{r_3}} = & 
                                    \begin{cases}
                                      \updReg{r_1}{\vsc} & 
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_2) = c \in \SealableCaps \text{ and } \\
                                        \Phi(r_3) = \seal{\sigma,\_} \text{ and } \\
                                        \vsc = \sealed{c,\sigma}
                                      \end{array}\\
                                      & \\
                                      \failed & \text{otherwise}
                                    \end{cases}
\end{align*}

>>>>

%%% Math notation
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\defbnf}{::=}
\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
\newcommand{\ssem}[2][\Phi]{\sem{#2}_{\mathrm{src}}(#1)}
\newcommand{\tsem}[2][\Phi]{\sem{#2}_{\mathrm{trg}}(#1)}
\newcommand{\dom}{\mathrm{dom}}

@@

% Text
\newcommand{\tand}{\text{ and }}
\newcommand{\totherwise}{\text{otherwise }}

@@

\newcommand{\fourinstr}[5]{
  \ifthenelse{\equal{#2#3#4#5}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2 \; #3 \; #4 \; #5}
}

@@

% One argument
\newcommand{\sjmp}[1]{\oneinstr{\src{jmp}}{#1}}
\newcommand{\spush}[1]{\oneinstr{\src{push}}{#1}}
\newcommand{\spop}[1]{\oneinstr{\src{pop}}{#1}}

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{isptr}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
\newcommand{\sgetl}[2]{\twoinstr{\src{getl}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
\newcommand{\scseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

% Four arguments
\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}

@@

% One argument
\newcommand{\tjmp}[1]{\oneinstr{\trg{jmp}}{#1}}

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{isptr}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
\newcommand{\tgetl}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}

% Three arguments
\newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
\newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
\newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
\newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}
\newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

% Four arguments
\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}

@@

%%% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\rn}{\var{rn}}
\newcommand{\reg}{\var{reg}}
\newcommand{\mem}{\var{mem}}
\newcommand{\ms}{\var{ms}}
\newcommand{\pc}{\var{pc}}
\newcommand{\stk}{\var{stk}}
\newcommand{\ret}{\var{ret}}
\newcommand{\data}{\var{data}}
\newcommand{\priv}{\var{priv}}
\newcommand{\opc}{\var{opc}}
\newcommand{\odata}{\var{odata}}
\newcommand{\vsc}{\var{sc}}
\newcommand{\cb}{\var{cb}}
\newcommand{\baddr}{\var{b}}
\newcommand{\eaddr}{\var{e}}
\newcommand{\aaddr}{\var{a}}
\newcommand{\stdrng}{[\baddr,\eaddr]}

@@

%%% Functions
\newcommand{\plainfun}[2]{
  \ifthenelse{\equal{#2}{}}
  {\mathit{#1}}
  {\mathit{#1}(#2)}
}
\newcommand{\encPerm}[1]{\plainfun{encPerm}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}

\begin{document}
\section{The two capability machines}
\subsection{Domains}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    & \aendbnf & \in & \Addr \uplus \{\infty \} \\
    &  \sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

In the source language, $\src{\Stack}$ is a call stack that contains the data for each call. The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) caller private stack, and 3) the address of the old stack pointer.

More convenient definitions
\[
  \begin{array}{rcl}
    \Phi(r) & \defeq & \Phi.\reg(r)
  \end{array}
\]
where $r\in \RegName$.

\subsection{Syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (Inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
\end{array}
\]

The source machine is also a capability machine with memory capabilities, local capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rret, \rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

\subsubsection{Helpful functions and sets}

\[
  \updPcAddr{} \defeq \dots
\]

@@

\[
  \readAllowed{} \defeq \dots
\]

\[
  \writeAllowed{} \defeq \dots
\]

\[
  \writeLocalAllowed{} \defeq \dots
\]

\[
  % Lau: this should work for everything that might be local - used in store.
  \isLoc{c} \defeq \dots
\]

\[
  \nonLoc{c} \defeq \dots
\]

@@

\[
  \opaquePerm{} \defeq \dots
\]

\[
  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rret, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
\]

@@

If we are not in one of the two cases above, then the execution fails.
\begin{align*}
  \ssem{\scall{r_1}{r_2}} = & 
    \begin{cases}
      (\mem,\reg',\stk',\ms_\stk') & \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        % "red conditions" (in reference to blackboard picture)
                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
                                        \nonExec{cb_2} \text{ and } \\
                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
                                        % "yellow conditions"
                                        \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
                                        \noRetStkReg{\reg} \text{ and } \\
                                        \noRetStkMs{\ms_\stk'} \text{ and } \\
                                        \nonLoc{\cb_1,\cb_2} \text{ and } \\
                                        \reg' = \reg''
                                        \begin{array}[t]{l}
                                        \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
                                        \update{\rretc}{\retptr}\\
                                        \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
                                        \end{array}
\\
                                        \text{where $\forall r \in \RegName \ldotp \nonLoc{\reg''(r)} \text{ and } \reg''(r) = \reg(r)$}\\
                                        \quad \text{$\text{ or } \reg''(r)=0 $}\\
                                        \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
                                        \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
                                        \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
                                      \end{array}\\
                                      & \\
      \undefined & \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        % "red conditions" (reference to blackboard picture)
                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
                                        \nonExec{cb_2} \text{ and } \\
                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
                                        % "yellow conditions"
                                        \text{not } \arraycolsep=0pt
                                        \left(\begin{array}{l}
                                          \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
                                          \noRetStkReg{\reg} \text{ and } \\
                                          \noRetStkMs{\ms_\stk'} \text{ and } \\
                                          \nonLoc{\cb_1,\cb_2}
                                        \end{array} \right)
                                      \end{array}\\
                                      & \\
      \failed & \text{otherwise}
    \end{cases}
\end{align*}

The $\scseal{}{}{}$ instruction seals a sealable capability with a seal. The base of the seal is used for the seal. The used seal only works if it still grants authority to seal (that is the base of the seal is smaller than or equal to the end).
\begin{align*}
  \ssem{\scseal{r_1}{r_2}{r_3}} = & 
                                    \begin{cases}
                                      \updReg{r_1}{\vsc} & 
                                      \arraycolsep=0pt
                                      \begin{array}[t]{l}
                                        \Phi(r_2) = c \in \SealableCaps \text{ and } \\
                                        \Phi(r_3) = \seal{\sigma,\_} \text{ and } \\
                                        \vsc = \sealed{c,\sigma}
                                      \end{array}\\
                                      & \\
                                      \failed & \text{otherwise}
                                    \end{cases}
\end{align*}

Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$. We cannot just use $\sjmp{}$ because it only takes one argument and we need to be able to jump to a code/data pair. To this end we introduce $\sxjmp{}{}$:
\begin{align*}
  \ssem{\sxjmp{r_1}{r_2}} = & 
                              \begin{cases}
                                \Phi
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  [\reg.\pcreg \mapsto c_1] \\
                                  [\reg.\rdata \mapsto c_2]
                                \end{array} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
                                  \nonExec{c_2}
                                \end{array}\\\\
%
                                (\mem,\reg',\stk',\ms_\stk') &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_1) = \retptrc \text{ and } \\
                                  \Phi(r_2) = \retptrd \text{ and } \\
                                  \Phi = (\mem, \reg, \stk, \ms_\stk) \text{ and } \\
                                  \stk = (\opc,\ms_\priv,\aaddr_\stk) :: \stk' \text{ and } \\
                                  \reg \arraycolsep=0pt
                                  \begin{array}[t]{l}
                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\\
                                  \end{array}
                                \end{array}\\\\
%
                                \failed & \text{otherwise}
                              \end{cases}
\end{align*}
\lau{questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}

Notice, $\rdata$ serves the same function as the IDC register in CHERI, i.e.\ this is where the data capability is put. In the interpretation of $\sxjmp{}{}$, we require $c_2$ to be non-executable because CHERI does it. We are, however, not certain whether this requirement is strictly necessary (maybe it is dangerous to $\sxjmp{r}{r}$ where $r$ points to the same capability? Maybe it can be use to unseal a code capability by first sealing another code capability with the seal, $\sxjmp{}{}$ with the newly sealed and the old sealed capability. Now both capabilities are unsealed). CHERI also checks whether $c_1$ is executable and in range, but we do this on every step of the operational semantics, so we do not repeat that check.
